Characteristic polynomials of p-adic matrices

### 2017-02-02 charpoly/charpoly.tex
Correction of 3.3.

The second ingredient we need is the reciprocal polynomial.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $P \in K[X]$ be of degree $d \leq n.$ 
We define its reciprocal polynomial as $P^{rec}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X]).$ We denote by $A^{rec}$ the matrix
whose coefficients are the reciprocal polynomials of the
coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec}=&com(X-M), \\
(\chi_M \times 1)^{rec}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}

@@

\noindent 9. \ \textbf{Return} $V^{rec}, \psi^{rec}.$

>>>>

The second ingredient we need is reciprocal polynomials.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $d \in \mathbb{N}$ and $P \in K[X]$ of degree at most $d.$ 
We define the reciprocal polynomial of order $f$ of $P$ as $P^{rec,d}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X])$ a matrix of polynomials of degree at most $d.$
We denote by $A^{rec,d}$ the matrix
whose coefficients are the reciprocal polynomials of order $d$ 
of the coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec,n-1}=&com(X-M), \\
(\chi_M \times 1)^{rec,n}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}
\begin{proof}
It all comes down to the following result:
let $A \in M_d(K[X])$ a matrix of polynomials of degree at most $1,$
then $det (A^{rec,1})=det(A)^{rec,d}.$
Indeed, one can use multilinearity of the determinant on $X^d det(A(1/X))$
to prove this result.
It directly implies the second part of the lemma.
As for the first part, since coefficients of the comatrix we are interested in
are determinants of a matrix of polynomials of degree at most 1, the result
is also clear.
\end{proof}

@@

\noindent 9. \ \textbf{Return} $V^{rec,n-1}, \psi^{rec,n}.$
